# Data transformation


```{r setup, include=FALSE}
 # keep this chunk in your .Rmd file
 knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r echo=FALSE}
library(dplyr)
library(jsonlite)
library(stringr)
library(ggplot2)
```

## read data ##

```
yelp_business_data <- jsonlite::stream_in(file("data/yelp_academic_dataset_business.json"))
```

```{r include = FALSE}
yelp_business_data <- jsonlite::stream_in(file("data/yelp_academic_dataset_business.json"))
```

## change formats ##

Extracted the time table and put it back to the original dataset.
Transformed all the character columns into factors except categories column.
Extracted the attributes table as a new dataset, and deleted the attributes table and opening hours table.

```
yelp_business_data$Monday <- yelp_business_data$hours$Monday...
yelp_business_data$city <- factor(yelp_business_data$city)...
yelp_business_attribute <- yelp_business_data[,12]
yelp_business_data <- yelp_business_data[,c(-12,-14)]
```

```{r}
yelp_business_data$Monday <- yelp_business_data$hours$Monday
yelp_business_data$Tuesday <- yelp_business_data$hours$Tuesday
yelp_business_data$Wednesday <- yelp_business_data$hours$Wednesday
yelp_business_data$Thursday <- yelp_business_data$hours$Thursday
yelp_business_data$Friday <- yelp_business_data$hours$Friday
yelp_business_data$Saturday <- yelp_business_data$hours$Saturday
yelp_business_data$Sunday <- yelp_business_data$hours$Sunday
yelp_business_data$city <- factor(yelp_business_data$city)
yelp_business_data$state <- factor(yelp_business_data$state)
yelp_business_data$postal_code <- factor(yelp_business_data$postal_code)
yelp_business_data$is_open <- factor(yelp_business_data$is_open)
yelp_business_attribute <- yelp_business_data[,12]
yelp_business_data <- yelp_business_data[,c(-12,-14)]
#yelp_business_data$categories <- factor(yelp_business_data$categories)
```

Extracted time information from character.
Changed the opening hours of one column to two separate columns in order to capture the data more easily.

```
yelp_business_data$monday1 <- str_extract(yelp_business_data$Monday,'^[0-9]+:[0-9]+')
yelp_business_data$monday2 <- str_extract(yelp_business_data$Monday,'[0-9]+:[0-9]+$')...
yelp_business_data <- yelp_business_data[,-c(13:19)]
```

```{r}
yelp_business_data$monday1 <- str_extract(yelp_business_data$Monday,'^[0-9]+:[0-9]+')
yelp_business_data$monday2 <- str_extract(yelp_business_data$Monday,'[0-9]+:[0-9]+$')
yelp_business_data$tuesday1 <- str_extract(yelp_business_data$Tuesday,'^[0-9]+:[0-9]+')
yelp_business_data$tuesday2 <- str_extract(yelp_business_data$Tuesday,'[0-9]+:[0-9]+$')
yelp_business_data$wednesday1 <- str_extract(yelp_business_data$Wednesday,'^[0-9]+:[0-9]+')
yelp_business_data$wednesday2 <- str_extract(yelp_business_data$Wednesday,'[0-9]+:[0-9]+$')
yelp_business_data$thursday1 <- str_extract(yelp_business_data$Thursday,'^[0-9]+:[0-9]+')
yelp_business_data$thursday2 <- str_extract(yelp_business_data$Thursday,'[0-9]+:[0-9]+$')
yelp_business_data$friday1 <- str_extract(yelp_business_data$Friday,'^[0-9]+:[0-9]+')
yelp_business_data$friday2 <- str_extract(yelp_business_data$Friday,'[0-9]+:[0-9]+$')
yelp_business_data$saturday1 <- str_extract(yelp_business_data$Saturday,'^[0-9]+:[0-9]+')
yelp_business_data$saturday2 <- str_extract(yelp_business_data$Saturday,'[0-9]+:[0-9]+$')
yelp_business_data$sunday1 <- str_extract(yelp_business_data$Sunday,'^[0-9]+:[0-9]+')
yelp_business_data$sunday2 <- str_extract(yelp_business_data$Sunday,'[0-9]+:[0-9]+$')
yelp_business_data <- yelp_business_data[,-c(13:19)]
```

Transformed the character into date type. 

```
yelp_business_data$monday1 <- strptime(yelp_business_data$monday1,'%H:%M')...
yelp_business_data$categories <- tolower(yelp_business_data$categories)

```
```{r}
yelp_business_data$monday1 <- strptime(yelp_business_data$monday1,'%H:%M')
yelp_business_data$monday2 <- strptime(yelp_business_data$monday2,'%H:%M')
yelp_business_data$tuesday1 <- strptime(yelp_business_data$tuesday1,'%H:%M')
yelp_business_data$tuesday2 <- strptime(yelp_business_data$tuesday2,'%H:%M')
yelp_business_data$wednesday1 <- strptime(yelp_business_data$wednesday1,'%H:%M')
yelp_business_data$wednesday2 <- strptime(yelp_business_data$wednesday2,'%H:%M')
yelp_business_data$thursday1 <- strptime(yelp_business_data$thursday1,'%H:%M')
yelp_business_data$thursday2 <- strptime(yelp_business_data$thursday2,'%H:%M')
yelp_business_data$friday1 <- strptime(yelp_business_data$friday1,'%H:%M')
yelp_business_data$friday2 <- strptime(yelp_business_data$friday2,'%H:%M')
yelp_business_data$saturday1 <- strptime(yelp_business_data$saturday1,'%H:%M')
yelp_business_data$saturday2 <- strptime(yelp_business_data$saturday2,'%H:%M')
yelp_business_data$sunday1 <- strptime(yelp_business_data$sunday1,'%H:%M')
yelp_business_data$sunday2 <- strptime(yelp_business_data$sunday2,'%H:%M')
yelp_business_data$categories <- tolower(yelp_business_data$categories)
```

```{r echo = FALSE}
#backup
yelp_business_data1 <- yelp_business_data
```

## Deal with categories ##

Filtered the dataset by applying if there is a 'restaurant' in the categories.
Deleted some rows with 'restaurant', but not what we want.

```
cate <- strsplit(yelp_business_data$categories,',')
match = 'restaurants'
boolean1 <- grepl(match,cate)
yelp_restaurant <- yelp_business_data[boolean1,]
```

```{r}
cate <- strsplit(yelp_business_data$categories,',')
match = 'restaurants'
boolean1 <- grepl(match,cate)
yelp_restaurant <- yelp_business_data[boolean1,]
```

```
cate <- strsplit(yelp_restaurant$categories,',')
match2 = 'shopping'
boolean2 <- grepl(match2,cate)
yelp_restaurant <- yelp_restaurant[!boolean2,]
```

```{r}
cate <- strsplit(yelp_restaurant$categories,',')
match2 = 'shopping'
boolean2 <- grepl(match2,cate)
yelp_restaurant <- yelp_restaurant[!boolean2,]
cate <- strsplit(yelp_restaurant$categories,',')
match3 = 'arts'
boolean3 <- grepl(match3,cate)
yelp_restaurant <- yelp_restaurant[!boolean3,]
cate <- strsplit(yelp_restaurant$categories,',')
match4 = 'home'
boolean4 <- grepl(match4,cate)
yelp_restaurant <- yelp_restaurant[!boolean4,]
cate <- strsplit(yelp_restaurant$categories,',')
match5 = 'repair'
boolean5 <- grepl(match5,cate)
yelp_restaurant <- yelp_restaurant[!boolean5,]
cate <- strsplit(yelp_restaurant$categories,',')
match6 = 'beauty'
boolean6 <- grepl(match6,cate)
yelp_restaurant <- yelp_restaurant[!boolean6,]
```

```
yelp_restaurant <- subset(yelp_restaurant,!is.na(yelp_restaurant$business_id))
yelp_restaurant$categories <- gsub('restaurants[,]*','',yelp_restaurant$categories)
yelp_restaurant$categories <- gsub('food[,]*','',yelp_restaurant$categories)
```

```{r}
yelp_restaurant <- subset(yelp_restaurant,!is.na(yelp_restaurant$business_id))
yelp_restaurant$categories <- gsub('restaurants[,]*','',yelp_restaurant$categories)
yelp_restaurant$categories <- gsub('food[,]*','',yelp_restaurant$categories)
```

```{r echo = FALSE}
#backup
yelp_restaurant1 <- yelp_restaurant
```

```{r echo = FALSE}
#summary(yelp_business_data$categories)
```

Extracted words from categories.

```
cate <- strsplit(yelp_restaurant$categories,',|&|[)]|[(]|[ ]|/')
yelp_restaurant1$categories <- substr(cate,2,1000)
```

```{r }
cate <- strsplit(yelp_restaurant$categories,',|&|[)]|[(]|[ ]|/')
yelp_restaurant1$categories <- substr(cate,2,1000)
```

```{r echo = FALSE}
#install.packages('jiebaR')
#install.packages('jiebaRD')
#install.packages('wordcloud2')
#install.packages('tm')
library(jiebaR)
library(jiebaRD)
library(wordcloud2)
library(tm)
```

Word cloud of cleaned categories.

```
text<-yelp_restaurant1$categories
mixseg<-worker("mix") 
a<-segment(text,mixseg)
stopwords <- read.table("stopwords.txt")
stopwords <- as.vector(stopwords[,1]) 
wordResult <- removeWords(a,stopwords)
freq<-table(wordResult)
wordcloud2(freq,size=5)
```

```{r}
text<-yelp_restaurant1$categories
```

```{r}
mixseg<-worker("mix") 
a<-segment(text,mixseg) 
```

```{r}
stopwords <- read.table("stopwords.txt")
stopwords <- as.vector(stopwords[,1]) 
wordResult <- removeWords(a,stopwords)
```

```{r}
freq<-table(wordResult)
wordcloud2(freq,size=5)
```

Deeper cleaning and merging of categories.

```
yelp_restaurant1$categories <- gsub('"veggies"|"vegetarian"','"vegan"',yelp_restaurant1$categories)...
yelp_restaurant1$categories <- gsub('[(]|["]|[)]|[,]','',yelp_restaurant1$categories)
```

```{r}
yelp_restaurant1$categories <- gsub('"veggies"|"vegetarian"','"vegan"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"middle"','"middleeastern"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"coffeeshops"|"cafes"|"coffee"|"beverage"|"cafeteria"|"cideries"|"juice"|"kombucha"|"tea"|"cafems"|"smoothies"|"cafe"|"soup"|"yogurt"','"drinks"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"beer"|"cocktail"|"champagne"|"wine"|"whiskey"|"distilleries"|"breweries"|"pubs"|"brewpubs"|"gastropub"|"gastropubs"|"pub"|"bars"|"izakaya"|"bar"|"brasseries"|"bistros"|"speakeasies"','"alcohol"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"churros"|"catalan"|"empanadas"','"spainish"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"pancakes"|"cakes"|"cake"|"candy"|"cupcakes"|"cream"|"creperies"|"popcorn"|"sugar"|"waffles"|"cheesescafeks"|"chocolatiers"|"donuts"|"chocolatiers"|"macarons"|"offdessertss"|"offdesserts"|"practdesserts"|"servdessertss"|"servdesserts"|"spdessertss"|"ice"','"desserts"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"bulgarian"|"eritrean"|"guamanian"|"iberian"|"mauritius"|"nicaraguan"|"polynesian"|"slovakian"|"scottish"|"tuscan"|"uzbek"|"trinidadian"|"calabrian"|"senegalese"|"sicilian"|"swiss"','"othercountries"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"rotisserie"|"donairs"|"kebab"|"roasteries"|"steakhouses"|"cheesesteaks"|"fish"|"chicken"|"poke"','"meat"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"bagels"|"bakeries"|"dumplings"|"pasta"|"ramen"|"noodles"|"udon"|"patisserie"','"flourfood"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"pretzels"','"german"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"rican"','"puerto"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"spainish"|"tapas"','"spanish"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"sushi"|"tempura"|"teppanyaki"','"japanese"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"burgers"|"chips"|"falafel"|"pizza"|"delicatessen"|"delis"|"pita"|"junk"','"fast"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"bowls"|"sandwiches"','"salad"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"fondue"|"pot"|"szechuan"|"dim"|"cantonese"|"taiwanese"','"chinese"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"tacos"','"mexican"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('"poutineries"','"french"',yelp_restaurant1$categories)
yelp_restaurant1$categories <- gsub('[(]|["]|[)]|[,]','',yelp_restaurant1$categories)
```

Extracted unique word for each row.

```
cate <- strsplit(yelp_restaurant1$categories,' ')
for (i in 1:nrow(yelp_restaurant1)) {yelp_restaurant1$categories[i] <- str_c(unique(cate[[i]]),collapse=' ')}
```

```{r}
cate <- strsplit(yelp_restaurant1$categories,' ')
for (i in 1:nrow(yelp_restaurant1)) {yelp_restaurant1$categories[i] <- str_c(unique(cate[[i]]),collapse=' ')}
```

Word cloud about cuisine.

```
text1<-yelp_restaurant1$categories
mixseg1<-worker("mix") 
b<-segment(text1,mixseg1) 
wordResult1 <- removeWords(b,stopwords)
new_list <- c('acai','alcohol','barbeque','breakfast','brunch','buffets','delivery','desserts','dinner','drinks','flourfood','fruits','imported','local','organic','salad','vegan','meat','kosher')
wordDraw <- removeWords(wordResult1,new_list)
freq1<-table(wordDraw)
wordcloud2(freq1,size=12)
```

```{r}
text1<-yelp_restaurant1$categories
mixseg1<-worker("mix") 
b<-segment(text1,mixseg1) 
```

```{r}
wordResult1 <- removeWords(b,stopwords)
new_list <- c('acai','alcohol','barbeque','breakfast','brunch','buffets','delivery','desserts','dinner','drinks','flourfood','fruits','imported','local','organic','salad','vegan','meat','kosher')
wordDraw <- removeWords(wordResult1,new_list)
```

```{r}
freq1<-table(wordDraw)
wordcloud2(freq1,size=12)
```

Cuisine countries.

```{r}
freq1
```

Delete rows without these categories.

```
wordResult1 = wordResult1[wordResult1 != '']
exist_cate <- str_c(wordResult1,collapse='|')
boolean_cate <- str_detect(yelp_restaurant1$categories,exist_cate)
yelp_restaurant1 <- yelp_restaurant1[boolean_cate,]
```

```{r}
wordResult1 = wordResult1[wordResult1 != '']
exist_cate <- str_c(wordResult1,collapse='|')
boolean_cate <- str_detect(yelp_restaurant1$categories,exist_cate)
yelp_restaurant1 <- yelp_restaurant1[boolean_cate,]
```

Added a new column for cuisine countries.

```
k <- names(freq1)
country <- str_c(k[-1],collapse='|')
yelp_restaurant1$country <- str_match_all(yelp_restaurant1$categories,country)
yelp_restaurant1$country <- gsub('c[(]+|[)]+|["]','',yelp_restaurant1$country)
yelp_restaurant1$country <- gsub('[", "]',' ',yelp_restaurant1$country)
yelp_restaurant1$country[yelp_restaurant1$country == 'character(0'] = NA
```

```{r}
k <- names(freq1)
country <- str_c(k[-1],collapse='|')
yelp_restaurant1$country <- str_match_all(yelp_restaurant1$categories,country)
yelp_restaurant1$country <- gsub('c[(]+|[)]+|["]','',yelp_restaurant1$country)
yelp_restaurant1$country <- gsub('[", "]',' ',yelp_restaurant1$country)
```

```{r}
yelp_restaurant1$country[yelp_restaurant1$country == 'character(0'] = NA
```


```{r echo = FALSE}
#category <- str_c(new_list,collapse='|')
```

Added new columns for categories.

```
yelp_restaurant1$acai <- ifelse(str_detect(yelp_restaurant1$categories,"acai"),1,0)...
```

```{r}
yelp_restaurant1$acai <- ifelse(str_detect(yelp_restaurant1$categories,"acai"),1,0)
yelp_restaurant1$alcohol <- ifelse(str_detect(yelp_restaurant1$categories,"alcohol"),1,0)
yelp_restaurant1$barbeque <- ifelse(str_detect(yelp_restaurant1$categories,"barbeque"),1,0)
yelp_restaurant1$breakfast <- ifelse(str_detect(yelp_restaurant1$categories,"breakfast"),1,0)
yelp_restaurant1$brunch <- ifelse(str_detect(yelp_restaurant1$categories,"brunch"),1,0)
yelp_restaurant1$buffets <- ifelse(str_detect(yelp_restaurant1$categories,"buffets"),1,0)
yelp_restaurant1$delivery <- ifelse(str_detect(yelp_restaurant1$categories,"delivery"),1,0)
yelp_restaurant1$desserts <- ifelse(str_detect(yelp_restaurant1$categories,"desserts"),1,0)
yelp_restaurant1$dinner <- ifelse(str_detect(yelp_restaurant1$categories,"dinner"),1,0)
yelp_restaurant1$drinks <- ifelse(str_detect(yelp_restaurant1$categories,"drinks"),1,0)
yelp_restaurant1$flourfood <- ifelse(str_detect(yelp_restaurant1$categories,"flourfood"),1,0)
yelp_restaurant1$fruits <- ifelse(str_detect(yelp_restaurant1$categories,"fruits"),1,0)
yelp_restaurant1$imported <- ifelse(str_detect(yelp_restaurant1$categories,"imported"),1,0)
yelp_restaurant1$local <- ifelse(str_detect(yelp_restaurant1$categories,"local"),1,0)
yelp_restaurant1$organic <- ifelse(str_detect(yelp_restaurant1$categories,"organic"),1,0)
yelp_restaurant1$salad <- ifelse(str_detect(yelp_restaurant1$categories,"salad"),1,0)
yelp_restaurant1$vegan <- ifelse(str_detect(yelp_restaurant1$categories,"vegan"),1,0)
yelp_restaurant1$meat <- ifelse(str_detect(yelp_restaurant1$categories,"meat"),1,0)
yelp_restaurant1$kosher <- ifelse(str_detect(yelp_restaurant1$categories,"kosher"),1,0)
```

```{r echo = FALSE}
#for (i in new_list) { yelp_restaurant1 <- data.frame(append(yelp_restaurant1,str_detect(yelp_restaurant1$categories,i) )) }
```

```
yelp_restaurant1 <- yelp_restaurant1[,-grep('categories',colnames(yelp_restaurant1))]
```

```{r}
yelp_restaurant1 <- yelp_restaurant1[,-grep('categories',colnames(yelp_restaurant1))]
```

## Output cleaned data ##

```
write.csv(x = yelp_restaurant1,file = "yelp_cleaned_data.csv")
```
```
write.csv(x = yelp_business_attribute,file = "yelp_attribute_data.csv")
```

```{r}
write.csv(x = yelp_restaurant1,file = "yelp_cleaned_data.csv")
```

```{r}
write.csv(x = yelp_business_attribute,file = "yelp_attribute_data.csv")
```


















